---
title: "DATA605_HW4"
author: "Gabriel Santos"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r}

library(jpeg)

library(RSpectra)
```


```{r}
files <- list.files(path = "E:\\Users\\tqkar\\Documents\\DATA605\\Week 4\\jpg", pattern = ".jpg", full.names = TRUE)
height=120; 
width=250;
fhi = matrix(1, nrow = 3, ncol = 3)
fhi[2, 2] = -8
plot_jpeg = function(path, add=FALSE)
{ jpg = readJPEG(path, native=T) # read the file
  res = dim(jpg)[2:1] # get the resolution, [x, y]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]),asp=1,type='n',xaxs='i',yaxs='i',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
  rasterImage(jpg,1,1,res[1],res[2])
}
im=array(rep(0,length(files)*height*width*3), dim=c(length(files), height, width,3))

for (i in 1:length(files)){

  temp= EBImage::resize(readJPEG(files[i]),120, 250)
  im[i,,,]=array(temp,dim=c(1, 120, 250,3))
}
flat=matrix(0, length(files), prod(dim(im)))
for (i in 1:length(files)){
  r=as.vector(im[i,,,1]); g=as.vector(im[i,,,2]);b=as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
shoes=as.data.frame(t(flat))

```


```{r}
par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))
for (i in 1:length(files)){ 
plot_jpeg(writeJPEG(im[i,,,]))
}

```


```{r}
scaled=scale(shoes, center = TRUE, scale = TRUE)
Sigma_=cor(scaled)
myeigen=eigs(Sigma_,5,which="LM")
cumsum(myeigen$values) / sum(eigen(Sigma_)$values)
```


#### *We see that even only using 5 EigenValues we are already well above 80% of the variability.*

```{r}
scaling=diag(myeigen$values[1:5]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes=scaled%*%myeigen$vectors[,1:5]%*%scaling
```

```{r}
newdata=im
dim(newdata)=c(length(files),height*width*3)
mypca=princomp(t(as.matrix(newdata)), scores=TRUE, cor=TRUE)
```


```{r}
pcaScores=t(mypca$scores)
dim(pcaScores)=c(length(files),height,width,3)
par(mfrow=c(5,5))
par(mai=c(.001,.001,.001,.001))
for (i in 1:length(files)){
plot_jpeg(writeJPEG(pcaScores[i,,,], bg="white"))  
}
```


#### *When analyzing the images, I was able to observe that we have very small details that do not allow us to correctly analyze the variability.*

#### *To see the images in a better way and see the real differences, we blur the images.*

```{r}
for (i in 1:length(files)){

  temp= EBImage::gblur(EBImage::resize(readJPEG(files[i]),120, 250),sigma = 8, radius=5)
  im[i,,,]=array(temp,dim=c(1, 120, 250,3))
}
flat=matrix(0, length(files), prod(dim(im)))
for (i in 1:length(files)){
  r=as.vector(im[i,,,1]); g=as.vector(im[i,,,2]);b=as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
shoes=as.data.frame(t(flat))
par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))
for (i in 1:length(files)){ 
plot_jpeg(writeJPEG(im[i,,,]))
}
```


```{r}
scaled=scale(shoes, center = TRUE, scale = TRUE)
Sigma_=cor(scaled)
myeigen=eigs(Sigma_,5,which="LM")
cumsum(myeigen$values) / sum(eigen(Sigma_)$values)
```


#### *In order to see the real differences, I erased the small details to improve the variability.*

```{r}
pcaScores=t(mypca$scores)
dim(pcaScores)=c(length(files),height,width,3)
par(mfrow=c(5,5))
par(mai=c(.001,.001,.001,.001))
for (i in 1:length(files)){
plot_jpeg(writeJPEG(pcaScores[i,,,], bg="white"))  
}
```


